%% witnessfinding.tex - Lower bounds for witness-finding algorithms
%%
%% Copyright 2014 Jeffrey Finkelstein.
%%
%% This LaTeX markup document is made available under the terms of the Creative
%% Commons Attribution-ShareAlike 4.0 International License,
%% https://creativecommons.org/licenses/by-sa/4.0/.
\documentclass{article}

\usepackage{amsmath}
\usepackage{amssymb}
%% This must come before hyperref.
\usepackage{amsthm}
%% This is strongly recommended by biblatex.
\usepackage[english]{babel}
\usepackage[backend=biber]{biblatex}
\usepackage[T1]{fontenc}
%% This must come before csquotes.
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
%% This is strongly recommended by biblatex.
\usepackage{csquotes}
%% This must come before hyperref.
\usepackage{thmtools}
%% This must come before complexity.
\usepackage{hyperref}
\usepackage{complexity}
\usepackage[firstpage]{draftwatermark}
\usepackage{microtype}
\usepackage{textcomp}
\usepackage{tikz}

\LoadMicrotypeFile{cmr}
\SetProtrusion
    [load=lmr-T1]
    {encoding=T1, family=lmr}
    {
      \textquotedblright = {,1000},
      \textquotedblleft = {1000,},
      {'} = {,1000},
      {[} = {1000,},
      {]} = {,1000},
      {,} = {,1000},
      {:} = {,1000},
      {;} = {,1000},
      {.} = {,1000}
    }

%% Set the ``work-in-progress'' watermark for the first page.
\SetWatermarkLightness{0.9}
\SetWatermarkText{Work-in-progress}
\SetWatermarkFontSize{3.5cm}

\hypersetup{pdftitle={Lower bounds for witness-finding algorithms}, pdfauthor={Jeffrey Finkelstein}}

\addbibresource{witnessfinding.bib}

\declaretheorem[numberwithin=section]{theorem}
\declaretheorem[numberlike=theorem]{lemma}
\declaretheorem[numberlike=theorem]{proposition}
\declaretheorem[numberlike=theorem]{corollary}
\declaretheorem[numberlike=theorem]{conjecture}
\declaretheorem[numberlike=theorem, style=definition]{definition}

%% Custom commands are declared here.
\newcommand{\email}[1]{\textlangle\href{mailto:#1}{\nolinkurl{#1}}\textrangle}
\newcommand{\todo}[1]{\textbf{TODO #1}}
\newcommand{\mc}{\mathcal}
\newcommand{\floor}[1]{\left\lfloor{#1}\right\rfloor}
\newcommand{\ceil}[1]{\left\lceil{#1}\right\rceil}
\DeclareMathOperator{\pow}{pow}
\newcommand{\mono}{\mathsf{mono}}
\newcommand{\direct}{\mathsf{direct}}
\newcommand{\all}{\mathsf{all}}

%% Redefine the footnote environment so it has no reference and no number.
\long\def\symbolfootnote#1{\begingroup%
\def\thefootnote{\fnsymbol{footnote}}\footnotetext{#1}\endgroup}

\title{Lower bounds for witness-finding algorithms}
\author{Jeffrey Finkelstein}

\begin{document}

\maketitle

\symbolfootnote{%
  Copyright 2014 Jeffrey~Finkelstein \email{jeffreyf@bu.edu}.

  This document is licensed under the Creative Commons Attribution-ShareAlike 4.0 International License, which is available at \mbox{\url{https://creativecommons.org/licenses/by-sa/4.0/}}.
  The \LaTeX{} markup that generated this document can be downloaded from its website at \mbox{\url{https://github.com/jfinkels/witnessfinding}}.
  The markup is distributed under the same license.
}

\section{Introduction}
%%
%% context
Can a procedure that decides whether a Boolean formula has a satisfying assignment help to find such an assignment?
The naïve adaptive ``search-to-decision'' reduction uses a linear number of (quite weak) queries.
%% need
Is there a lower bound on the number of queries required for a nonadaptive search-to-decision reduction?
%% task
We report on lower bounds for various classes of queries.
%% object

%% findings
Most interesting types of queries require at least $\Omega(n^2)$ queries to find a witness.
%% conclusion
We were hoping to use lower bounds on these types of witness-finding algorithms to show that randomized reductions to $\CeP$ are less powerful than randomized reductions to $\PP$, but we were unable to do so.
%% perspectives

\autoref{fig:queries} contains a summary of the upper and lower bounds for nonadaptive witness-finding algorithms.
In addition, adaptive $\mc{Q}_0$ algorithms use $\Theta(n)$ queries, and adaptive $\mc{Q}_1$ witness-finding algorithms with high success probability use $\Theta(n^2)$ queries.
\todo{Currently the lower bound for the query complexity of adaptive $\mc{Q}_1$ witness-finding algorithms with success probability $\frac{1}{2}$ is $\Omega(n)$.
  Can we either improve this to $\Omega(n^2)$ or provide a $O(n)$ upper bound?
}

\todo{Compare these randomized witness-finding algorithms with ``bounded query hierarchy'' algorithms, in which the query generator and the witness generator are not randomized (but possibly nondeterministic), and the goal is to decide a language, not necessarily output a witness.}

\begin{figure}
  \caption{\label{fig:queries}%
    When considering nonadaptive witness-finding algorithms, $\mc{Q}_{\direct}$ algorithms require $\Omega(2^n)$ queries, $\mc{Q}_{\mono}$, $\mc{Q}_{\NP}$, and $\mc{Q}_k$ algorithms require $\Omega(n^2)$ queries, and $\mc{Q}_{\all}$ algorithms require $\Omega(n)$ queries.
    In this diagram, lower bounds propagate downward and upper bounds propagate upward.%
  }
  \begin{center}
    \begin{tikzpicture}[->, every node/.style={align=center}]
      \node at (0, 0) (direct) {$\mc{Q}_{\direct}$ \\ $\Theta(2^n)$};
      \node at (0, 2) (equal0) {$\mc{Q}_0$ \\ $\Theta(n^2)$};
      \node at (2, 2) (equal1) {$\mc{Q}_1$ \\ $\Theta(n^2)$};
      \node at (0, 5) (NP) {$\mc{Q}_{\NP}$ \\ $\Theta(n^2)$};
      \node at (-2, 4) (geqk) {$\cup_k \mc{Q}_{\geq k}$ \\ $\Theta(n^2)$};
      \node at (2, 4) (eqk) {$\cup_k \mc{Q}_k$ \\ $\Theta(n^2)$};
      \node at (-2, 6) (PP) {$\mc{Q}_{\PP}$ \\ $\Omega(n), O(n^2)$};
      \node at (2, 6) (CeP) {$\mc{Q}_{\CeP}$ \\ $\Omega(n), O(n^2)$};
      \node at (-4, 6) (mono) {$\mc{Q}_{\mono}$ \\ $\Theta(n^2)$};
      \node at (0, 7) (all) {$\mc{Q}_{\all}$ \\ $\Theta(n)$};
      \draw (direct) to (equal0);
      \draw (direct) to (equal1);
      \draw (equal0) to (NP);
      \draw (equal0) to (geqk);
      \draw (equal0) to (eqk);
      \draw (equal1) to (eqk);
      \draw (eqk) to (CeP);
      \draw (geqk) to (PP);
      \draw (geqk) to (mono);
      \draw (NP) to (CeP);
      \draw (NP) to (PP);
      \draw (PP) to (all);
      \draw (mono) to (all);
      \draw (CeP) to (all);

      \node[color=gray] at (-6, 0) (weak) {weaker queries};
      \node[color=gray] at (-6, 7) (strong) {stronger queries};
      \draw[color=gray] (weak) to (strong);
    \end{tikzpicture}
  \end{center}
\end{figure}

\section{Definitions}

Let $\Sigma = \{0, 1\}$.
For any natural number $n$, the set $\Sigma^n$ denotes the set of binary strings of length $n$.
For each $b \in \{0, 1\}$, let $\bar{b} = 1 - b$.
Let $[n] = \{1, \dotsc, n\}$.
Denote by $\log n$ the base-$2$ logarithm.

A language $L \subseteq \Sigma^*$ is a \emph{witnessed language} if there exists a language $L' \subseteq \Sigma^* \times \Sigma^*$, called the \emph{witness relation}, such that for each $x$ of length $n$, we have $x \in L$ if and only if there is a $w \in \Sigma^n$ such that $(x, w) \in L'$.
(We don't consider strings $w$ of length polynomial in $n$ because a language in any ``well-behaved'' complexity class can be converted to an equivalent language in which all witnesses have length exactly $n$ by padding the string $x$ by an appropriate amount.)

Suppose $n$ is a positive integer.
A \emph{witness set} is a nonempty subset of $\Sigma^n$ denoted by $W$, and each element of $W$ is a \emph{witness}.
A \emph{query} is a Boolean predicate $Q \colon \pow(\Sigma^n) \to \{0, 1\}$.

\begin{definition}[{\autocite{krw14}}]
  Let $\mc{W}$ be a collection of witness sets and $\mc{Q}$ a collection of queries.
  A \emph{(nonadaptive) $\mc{Q}$-query witness-finding algorithm} is a joint distribution $(Q_1, \dotsc, Q_m)$ on $\mc{Q}^m$, for some natural number $m$, and a function $f \colon \{0, 1\}^m \to \Sigma^n$ such that $\Pr[f(Q_1(W), \dotsc, Q_m(W)) \in W] > \frac{1}{2}$ for each $W \in \mc{W}$.
  The number $m$ is the \emph{query complexity} of the algorithm.
\end{definition}

This definition can be generalized to allow for success probabilities other than $\frac{1}{2}$, and to allow for adaptive queries.
Unless otherwise specified, the success probability is assumed to be $\frac{1}{2}$ and the queries are assumed to be nonadaptive.

\todo{Define an ``adaptive distribution sequence'' (which means that $Q_{i + 1}(W)$ is actually $Q_{i + 1}(W, Q_1(W), \dotsc, Q_{i - 1}(W))$).}

\begin{itemize}
\item
  An \emph{$(= k)$ query} is a query $Q_S$ of the form $Q_{S}(W) = 1$ if and only $|S \cap W| = k$, for any $S \subseteq \Sigma^n$ and any $W \in \pow(\Sigma^n)$.
  The collection of all $(= k)$ queries is denoted $\mc{Q}_k$.
\item
  A \emph{$(\geq k)$ query} is a query $Q_S$ of the form $Q_{S}(W) = 1$ if and only $|S \cap W| \geq k$, for any $S \subseteq \Sigma^n$ and any $W \in \pow(\Sigma^n)$.
  The collection of all $(\geq k)$ queries is denoted $\mc{Q}_{\geq k}$.
\item
  An element of $\mc{Q}_1$ is called an \emph{isolation query}.
\item
  An element of $\mc{Q}_0$ is called an \emph{emptiness query}.
\end{itemize}

\begin{definition}
  Let $\mc{W}$ be a collection of witness sets.
  A \emph{hitting algorithm} is an adaptive sequence of distributions $(Q_1, \dotsc, Q_m)$ on $\mc{Q}^m_1$, for some natural number $m$, such that $\Pr[Q_m(W) = 1] > \frac{1}{2}$.
  The number $m$ is the \emph{query complexity} of the algorithm.
\end{definition}

This definition is a complexity theoretic restatement of the combinatorial ``hitting game'', as described in \autocite[Section~3]{newport14} (the original definition seems to be from \autocite[Definition~5]{bgi92}).

A hitting algorithm is always an adaptive algorithm; a nonadaptive hitting algorithm makes little sense here because it would be essentially the same as simply guessing an element of the witness set at random.

A witness-finding algorithm must output an element of $W$, whereas a hitting algorithm need not.

\todo{define polynomial-time algorithms (provide polysize circuits that encode the query sets)}

\section{Relationship among algorithms}

\begin{proposition}\label{prop:flip}
  Let $k$ be a positive integer.
  \begin{itemize}
  \item There is a $\mc{Q}_{\geq k}$ witness-finding algorithm if and only if there is a $\mc{Q}_{< k}$ witness-finding algorithm.
  \item There is a $\mc{Q}_k$ witness-finding algorithm if and only if there is a $\mc{Q}_{\neq k}$ witness-finding algorithm.
  \end{itemize}
  Furthermore, in both cases, the query complexity of the algorithms is exactly the same.
\end{proposition}
\begin{proof}
  We prove the first equivalence; the second is similar.
  For each $Q_S$ in $\mc{Q}_{\geq k}$ there is a corresponding $\overline{Q_S}$ in $\mc{Q}_{< k}$ such that $\overline{Q_S(W)} = \overline{Q_S}(W)$ for each witness set $W$, and furthermore, this correspondence is a bijection.
  Therefore if we define the function $\bar{f}$ by
  \begin{equation*}
    \bar{f}(A_1, \dotsc, A_m) = f(\overline{A_1}, \dotsc, \overline{A_m}),
  \end{equation*}
  for all possible bits $A_1, \dotsc, A_m$, we get the desired equivalence: $((Q_1, \dotsc, Q_m), f)$ is equivalent to $((Q_1, \dotsc, Q_m), \bar{f})$.
\end{proof}

\begin{corollary}\label{cor:flip}
  There is a $\mc{Q}_0$ witness-finding algorithm if and only if there is a $\mc{Q}_{\neq 0}$ witness-finding algorithm.
\end{corollary}

\begin{proposition}\label{prop:leqgeq}
  Let $k$ and $m$ be natural numbers.
  If there is a $\mc{Q}_k$ witness-finding algorithm with query complexity $m$, then there is a $\mc{Q}_{\geq k} \cup \mc{Q}_{\geq k + 1}$ witness-finding algorithm with query complexity $2 m$.
\end{proposition}
\begin{proof}
  Suppose $((Q_{S_1}, \dotsc, Q_{S_m}), f)$ is the $\mc{Q}_k$ witness-finding algorithm.
  By \autoref{prop:flip}, it suffices to show a $\mc{Q}_{\geq k} \cup \mc{Q}_{\leq k}$ witness-finding algorithm.
  For each $Q_{S_i}$, construct two queries $Q^\geq_{S_i}$ and $Q^\leq_{S_i}$ defined as $Q^\geq_{S_i}(W) = 1$ if and only $|S_i \cap W| \geq k$ and $Q^\leq_{S_i}(W) = 1$ if and only if $|S_i \cap W| \leq k$.
  Define $f'$ by
  \begin{equation*}
    f'(A_1, \dotsc, A_m, B_1, \dotsc, B_m) = f(A_1 \land B_1, \dotsc, A_m \land B_m),
  \end{equation*}
  for all bits $A_1, \dotsc, A_m$ and all bits $B_1, \dotsc, B_m$.
  Since $Q^\geq_{S_i}(W) \land Q^\leq_{S_i}(W) = Q_{S_i}(W)$, the correctness of $((Q^\geq_{S_1}, \dotsc Q^\geq_{S_m}, Q^\leq_{S_1}, \dotsc, Q^\leq_{S_m}), f')$ follows from the correctness of $((Q_1, \dotsc, Q_m), f)$.
\end{proof}

The problem of finding a hitting algorithm reduces to the problem of finding a $\mc{Q}_1$ witness-finding algorithm.

\begin{lemma}\label{lem:reduction}
  Let $m$ be a natural number.
  If there is a $\mc{Q}_1$ witness-finding algorithm with query complexity $m$, then there is a hitting algorithm with query complexity $m + 1$.
\end{lemma}
\begin{proof}
  Suppose $((Q_{S_1}, \dotsc, Q_{S_m}), f)$ is the $\mc{Q}_1$ witness-finding algorithm.
  For any witness set $W$, let $S_{m + 1}$ be the singleton set containing the element $f(Q_{S_1}, \dotsc, Q_{S_m})$.
  Then $(Q_{S_1}, \dotsc, Q_{S_{m + 1}})$ is a hitting algorithm, since the probability that $f(Q_{S_1}(W), \dotsc, Q_{S_m}(W)) \in W$ is exactly the probability that $Q_{S_{m + 1}}(W) = 1$.
\end{proof}

On the other hand, there does not seem to be a general reduction from a $\mc{Q}_1$ witness-finding algorithm to a hitting algorithm.
In order to output a single witness knowing only that a witness is contained in a particular query set doesn't help, unless the query set is small; there is no such guarantee on the size of that set.

\begin{conjecture}
  Under an appropriate setting of parameters, if the existence of a hitting algorithm implies the existence of an isolation query algorithm, then $\RP = \NP$.
  \todo{But both types of algorithms exist with the same parameters, as shown in the next section.}
\end{conjecture}

Now, consider the possibility that a $\mc{Q}_k$ witness-finding algorithm implies a $\mc{Q}_{k - 1}$ witness-finding algorithm.
Intuitively, in order for this to work, the latter algorithm would have to simulate the queries $Q_i$ from the former algorithm with some sort of collection of queries of the form $Q_i \setminus \{q_{i, j}\}$, where $q_{i, j}$ is the $j$th member of $Q_i$.
However, the oracle answers would be inconsistent, because we don't know beforehand whether $q_{i, j}$ is in the witness set or not (causing a discrepancy between the answers of the two oracles).
A similar problem arises when attempting the converse simulation: the simulated queries would be something of the form $Q_i \cup \{v_j\}$, where $v_j$ is the $j$th element of $\Sigma^n \setminus Q_i$, thus there would be a discrepancy between the answers of the two oracles.
We therefore conjecture that there is evidence that such a simulation is unlikely.

\begin{conjecture}
  For each $k \in \left[2^n\right]$, if the existence of a $\mc{Q}_k$ witness-finding algorithm implies the existence of a $\mc{Q}_{k + 1}$ witness-finding algorithm, then $\RP = \NP$.
  \todo{But both such algorithms exist unconditionally, as shown in the next section.}
\end{conjecture}

%% \subsection{Reductions to counting classes}

%% In \autocite[Section~7]{green93}, Fred asks, ``Does the hypothesis $\P^{\CeP} = \P^\PP$ have any dramatic consequences?''
%% Our intuition is that $\P^{\CeP}$ is less powerful than $\P^\PP$, because the latter can determine the exact number of accepting paths in a computation (or in other words, the number of witnesses in a witness set) by a binary search, whereas the former does not seem to admit this possibility.
%% On the contrary, we know that $\NP^{\CeP} = \NP^\PP$.
%% Perhaps a first step toward comparing $\P^\PP$ and $\P^{\CeP}$ is comparing $\RP^\PP$ and $\RP^{\CeP}$.

%% There is a superficial similarity to the question ``is $\RP^\PP$ a subset of $\RP^{\CeP}$?'' and the question ``does a witness-finding algorithm with oracle $A^{\geq k}_W$ imply a witness-finding algorithm with oracle $A^{= k}_W$?''
%% However, we run into a few problems when trying to follow this approach.
%% First, according to \autocite[Remark~2]{krw12}, the lower bounds on emptiness query algorithms (with oracle $A^{=0}_W$) ``essentially hold'' when the oracle is $A^{\geq k}_W$ instead, for any $k$ bounded by a polynomial in $n$.
%% Furthermore, if we attempt such a proof, we run into several problems.

%% \begin{conjecture}
%%   If $\RP^\PP \subseteq \RP^{\CeP}$, then a witness-finding algorithm with oracle $A^{\geq k}_W$ implies a witness-finding algorithm with oracle $A^{= k}_W$.
%% \end{conjecture}
%% \begin{proof}[Broken proof]
%%   Let $W$ be any witness set.
%%   Let $(\mc{Q}, \mc{F})$ be a witness-finding algorithm with oracle $A^{= k}_W$.
%%   Let $L$ be a witnessed language with witness relation $L'$.
%%   The problem of deciding whether $|Q \cap W| \geq k$ given any query set $Q$ can be decided by a $\PP$ oracle.
%%   Construct machine $M$ with access to that oracle as follows on random string $s$ of length $r$.
%%   \begin{itemize}
%%   \item Run $\mc{Q}(s)$ to generate queries $Q_1, \dotsc, Q_m$, simulating access to the $\PP$ oracle (acting as the $A^{\geq k}_W$ oracle) when necessary.
%%   \item Get query answers $A_1, \dotsc, A_m$ from the oracle.
%%   \item Run $\mc{F}(s, A_1, \dotsc, A_m)$ to get potential witness $w$.
%%   \item Decide if $(x, w) \in L'$.
%%   \end{itemize}
%%   \todo{Problem 1: need to show that this is a correct $\RP^\PP$ algorithm.}
%%   By our hypothesis that $\RP^\PP \subseteq \RP^{\CeP}$, there is also a randomized oracle machine $M'$ that decides $L$ with access to a $\CeP$ oracle.
%%   Now we wish to use $M'$ to construct a witness-finding algorithm $(\mc{Q}', \mc{F}')$ with access to oracle $A^{= k}_W$.

%%   Define $\mc{Q}'$ as follows on random string $s$ of length $r$.
%%   \begin{itemize}
%%   \item Run $M'(x, s)$ and intercept each query $q_1, \dotsc, q_{m(n)}$ that it would make to the $\CeP$ oracle.
%%     \todo{Problem 2: $M'$ needs access to a particular input string $x$; we may be able to modify the definition of the witness-finding algorithms to allow this.}
%%     Each query $q_i$ is a question of the form ``is $|W_{q_i}| = k$?'', where $W_{q_i}$ is a witness set that corresponds to the input string $q_i$, so we simulate those queries by sending the query set $W_{q_i}$ to the oracle $A^{= k}_W$.
%%     \todo{Problem 3: this won't work, since $A^{= k}_W$ only answers questions of the form ``is $|W_{q_i} \cap W| = k$?'', but we need it to answer questions like ``is $|W_{q_i}| = k$?'', and we don't get to choose the arbitrary witness set $W$.}
%%   \item Output the queries $W_{q_1}, \dotsc, W_{q_{m(n)}}$.
%%   \end{itemize}
%%   Define $\mc{F}'$ as follows on random string $s$ and oracle answers $A_1, \dotsc, A_{m(n)}$.
%%   \begin{itemize}
%%   \item Run $M'(x, s)$ and when $M'$ makes query $q_i$, provide answer $A_i$ to it.
%%   \item \todo{Problem 4: at this point, all $M'$ tells us is that there is a witness (or none); it gives no indication as to what the witness is.
%%     If one of the queries causes the oracle to answer yes and the size of the query set is not too large, we could simply output a random element of the query set for a small penalty in success probability, but in general we have no such guarantees.}
%%   \end{itemize}
%% \end{proof}

%% Although this starts to have little to do with the witness-finding algorithm results given in the following sections, another approach to showing that $\P^{\CeP} \neq \P^\PP$ may be to show that adaptive queries to $\CeP$ are no more powerful than nonadaptive queries, whereas adaptive queries to $\PP$ are more powerful than adaptive queries.
%% What is known about $\PP$?
%% \begin{itemize}
%% \item $\PP$ is closed under complement, union, and intersection.
%% \item $\PP = \P^{\PP[O(1)]}_\parallel$ (that is, $\PP$ is closed under polynomial time $k$-round truth-table reductions, for any $k \in O(1)$).
%% \end{itemize}
%% What is known about $\CeP$?
%% \begin{itemize}
%% \item $\P^{\CeP[O(1)]}_\parallel = \P^{\CeP[O(\log n)]}$ (see \autocite[Theorem~5]{green93} and \autocite[Corollary~4.6]{ogiwara94}).
%% \end{itemize}

\section{Upper bounds}

\begin{theorem}\label{thm:naive}
  There is a polynomial time adaptive $\mc{Q}_0$ witness-finding algorithm with success probability $1$ and query complexity $O(n)$.
\end{theorem}
\begin{proof}[Proof idea]
  By \autoref{cor:flip}, it suffices to show a $\mc{Q}_{\neq 0}$ witness-finding algorithm.
  This the well-known bit-by-bit $\P^\NP$ computable function for computing a witness for a given string.

  For each $i \in [n]$ and each $b \in \{0, 1\}$, define the query $S_{i, b}$ inductively (or in other words, adaptively) as follows.
  If $i = 1$, then $S_{i, b} = \{ b v \, | \, v \in {\{0, 1\}}^{n - 1}\}$.
  For each subsequent pair of queries, let $u = b_1 \dotsb b_{i - 1}$, where $b_j$ is a bit that caused the oracle to answer ``yes'' on query $S_{j, b_j}$ for each $j \in [i - 1]$, and let $S_{i, b} = \{ u b v \, | \, v \in {\{0, 1\}}^{n - i}\}$.
  Finally, given the responses to $S_{n, 0}$ and $S_{n, 1}$, which are singleton sets, the algorithm simply outputs the sole element of the query set that caused the oracle to answer ``yes''.

  The total number of queries is $2n$, which is in $O(n)$.
\end{proof}

This upper bound means that any class of queries containing $\mc{Q}_0$ as a subset also satisfies this very strong upper bound.

\begin{theorem}[{\autocite[Theorem~1.4]{krw14} (citing \autocite{bcgl92})}]\label{thm:vv}
  There is a polynomial time $\mc{Q}_0$ witness-finding algorithm with query complexity $O(n^2)$.
\end{theorem}
\begin{proof}[Proof idea]
  By \autoref{cor:flip}, it suffices to show a $\mc{Q}_{\neq 0}$ witness-finding algorithm.
  This algorithm uses the Valiant--Vazirani Isolation Lemma in each query to guess each bit of a possible witness (in parallel).

  First, consider a simpler set of queries.
  Let $s$ be a string in $\Sigma^r$.
  For each $i \in [n]$ and each $b \in \Sigma$, let $S_{i, b} = \{ w \, | \, w \in C_s \land w_i = b\}$, where $C$ is the isolating set from the Valiant--Vazirani Isolation Lemma (for example, the set of all strings whose image is zero under a pairwise independent hash function).
  The Valiant--Vazirani Isolation Lemma proves that this algorithm has success probability $\Omega(\frac{1}{n})$ and query complexity $O(n)$.
  In order to get success probability $\Omega(1)$, we create $O(n)$ collections of $O(n)$ queries, with an independent set $C_{s_k}$ for each collection.
  Now our queries are $S_{k, i, b} = \{ w \, | \, w \in C_{s_k} \land w_i = b\}$, for each $k \in \left[O(n)\right]$, each $i$, and each $b$.
\end{proof}

In \autocite{krw12}, the authors call the above theorem the ``truth-table reduction'' analog of the ``many-one reduction'' result of \autocite[Theorem~4.2]{dkvmw13}.
However, this is not strictly precise, since the former involves $\mc{Q}_0$ queries, whereas the latter involves $\mc{Q}_1$ queries.
Fortunately, the proof uses the Valiant--Vazirani Isolation Lemma (that is, the sets $C_{s_k}$), so the probability that $|S \cap W| \neq 0$ equals the probability that $|S \cap W| = 1$.
(On the other hand, there is no other known relationship between $\mc{Q}_0$ witness-finding algorithms and $\mc{Q}_1$ witness-finding algorithms, other than the fact that $\mc{Q}_0 \cup \mc{Q}_1 \subseteq \cup_{k \in \mathbb{N}} \mc{Q}_k$.)

\begin{corollary}\label{cor:q1upper}
  There is a polynomial time $\mc{Q}_1$ witness-finding algorithm with query complexity $O(n^2)$.
\end{corollary}

\todo{Double check this logic.}

\begin{corollary}
  There is a polynomial time hitting algorithm with query complexity $O(n^2)$.
\end{corollary}
\begin{proof}
  Combine the previous corollary with \autoref{lem:reduction}.
\end{proof}

\section{Lower bounds}

A query $Q$ is a \emph{monotone query} if for all $W$ and $W'$, subsets of $\Sigma^n$, if $W \subseteq W'$, then $Q(W) = 1$ implies $Q(W') = 1$.
The collection of all monotone queries is denoted $\mc{Q}_{\mono}$.

\begin{theorem}[{\autocite[Theorem~1.3]{krw14}}]\label{thm:monotonelower}
  Each $\mc{Q}_{\mono}$ witness-finding algorithm has query complexity $\Omega(n^2)$.
\end{theorem}

The proof of this theorem relies on the existence of a distribution on witness sets that places all witnesses far apart, so it is hard to guess both the number and location of the witnesses.
This distribution comes from \autocite[Theorem~4.2]{dkvmw13}.

An infinite sequence of queries $(Q^1, Q^2, \dotsc)$ is an \emph{$\NP$ query} if there is a nondeterministic oracle Turing machine $M$ such that for each $n$, we have $Q^n(W) = 1$ if and only if $M^W(1^n)$ accepts (where $Q^n$ denotes a query on witness sets of strings of size $n$).
The collection of all $\NP$ queries is denoted $\mc{Q}_{\NP}$.

\begin{theorem}[{\autocite[Theorem~1.6]{krw14}}]\label{thm:nplower}
  Each $\mc{Q}_{\NP}$ witness-finding algorithm has query complexity $\Omega(n^2)$.
\end{theorem}

\begin{corollary}[{\autocite[Theorem~2]{krw12}}]\label{cor:q0lower}
  Each $\mc{Q}_0$ witness-finding algorithm has query complexity $\Omega(n^2)$.
\end{corollary}
\begin{proof}
  This is a special case of both \autoref{thm:monotonelower} and \autoref{thm:nplower}.
\end{proof}

The next corollary is a generalization of \autocite[Remark~2]{krw12}.

\begin{corollary}
  Suppose $k \in \left[2^n\right]$.
  Each $\mc{Q}_{\geq k}$ witness-finding algorithm has query complexity $\Omega(n^2)$.
\end{corollary}
\begin{proof}
  This is a special case of \autoref{thm:monotonelower}.
\end{proof}

\begin{corollary}
  Suppose $k \in \left[2^n\right]$.
  Each $\mc{Q}_k$ witness-finding algorithm has query complexity $\Omega(n^2)$.
\end{corollary}
\begin{proof}
  This is a proof by contradiction.
  Assume there is a $\mc{Q}_k$ witness-finding algorithm with query complexity $o(n^2)$.
  By \autoref{prop:leqgeq}, there is a $\mc{Q}_{\geq k} \cup \mc{Q}_{\geq k + 1}$ witness-finding algorithm with query complexity $o(n^2)$.
  Since $\mc{Q}_{\geq k}$ and $\mc{Q}_{\geq k + 1}$ are both subsets of the collection of all monotone queries, each $\mc{Q}_{\geq k} \cup \mc{Q}_{\geq k + 1}$ witness-finding algorithm has query complexity $\Omega(n^2)$, by \autoref{thm:monotonelower}.
  This is a contradiction.
\end{proof}

\begin{corollary}\label{cor:q1lower}
  Each $\mc{Q}_1$ witness-finding algorithm has query complexity $\Omega(n^2)$.
\end{corollary}
\begin{proof}
  This is a special case of the previous corollary.
\end{proof}

\begin{theorem}[{\autocite[Theorems~3.3 and 3.4]{newport14}}]
  Each hitting algorithm with success probability $\Omega(1 - 2^{-n})$ has query complexity $\Omega(n^2)$.
  Furthermore, this is true even when restricted to witness sets of cardinality two.
\end{theorem}

The proof of this theorem relies on the probabilistic method to show the existence of a witness set that is hard to ``hit'', found in \autocite{ablp91}.

\begin{corollary}\label{cor:isolationalg}
  Each adaptive $\mc{Q}_1$ isolation query algorithm with success probability $\Omega(1 - 2^{-n})$ has query complexity $\Omega(n^2)$.
  Furthermore, this is true even when restricted to witness sets of cardinality two.
\end{corollary}
\begin{proof}
  This is a proof by contradiction.
  Assume there is an adaptive $\mc{Q}_1$ witness-finding algorithm with success probability $\Omega(1 - 2^{-n})$ and query complexity $o(n^2)$.
  By \autoref{lem:reduction}, there is a hitting algorithm with the same success probability and query complexity (up to constant factors).
  This would contradict the previous theorem, hence no such algorithm exists.
\end{proof}

%% We know that for adaptive $\mc{Q}_0$ witness-finding algorithms, there is a query complexity upper bound of $O(n)$ (see \autoref{thm:naive}), matching the lower bound of $\Omega(n)$.
%% Since we don't know whether $\mc{Q}_0$ witness-finding algorithms are comparable to $\mc{Q}_1$ witness-finding algorithms, we can't apply upper bounds for the former to the latter.
%% The only upper bound for adaptive $\mc{Q}_1$ witness-finding algorithms is the $O(n^2)$ upper bound for nonadaptive algorithms, \autoref{cor:q1upper}.
%% So it could be the case that there is an adaptive $\mc{Q}_1$ witness-finding algorithms admi

\begin{theorem}[{\autocite[Theorem~3.2]{newport14}}]
  Each hitting algorithm with expected success probability ??? has expected query complexity $\Omega(n)$.
\end{theorem}

\todo{I can't understand what the random variable is for which there is an expectation here.}

\section{Entropy on bad witness distributions}

In this section $H(X)$ denotes the Shannon entropy of random variable $X$ chosen from a distribution on binary strings, and $H_b(p)$ denotes the binary entropy function on a Bernoulli process with success probability $p$.

Define $\mc{W}_n^*$, a distribution on witness sets (that is, subsets of $\Sigma^n$), by the following sampling procedure.
\begin{itemize}
\item Let $W = \emptyset$.
\item Choose $D$ uniformly at random from $[n]$.
\item For each $w \in \Sigma^n$, add $w$ to $W$ with probability $2^{-D}$.
\item Output $W$.
\end{itemize}
(Technically, $W$ may be the empty set, but this occurs with probability $o(1)$, so for our purposes, we ignore it.
\todo{a brief explanation can be found in \autocite{krw14}, I believe.})

\begin{lemma}\label{lem:prob}
  For any positive integer $n$, for any $k \in [2^n]$, for any $d \in [n]$, and for any $S \subseteq \left(\left\{0, 1\right\}^n \setminus \emptyset\right)$,
  \begin{equation*}
    \Pr_{W \gets \mc{W}_n^*}\left[ |S \cap W| = k \middle| D = d\right] < \frac{e^{\frac{1}{12s}}}{\sqrt{2 \pi}} \left(\frac{s}{k (s - k)}\right)^{\frac{1}{2}},
  \end{equation*}
  where $s$ denotes the cardinality of $S$.
\end{lemma}
\begin{proof}
  For the sake of brevity, let $s$ denote $|S|$, and let $p$ denote the probability $\Pr_{W \gets \mc{W}_n^*}[ |S \cap W| = k | D = d]$.
  Without loss of generality, assume $s \geq k$, since if $s$ is smaller than $k$, the probability $p$ is zero.
  Since each element in $W$ is chosen independently with probability $2^{-d}$, the probability that $|S \cap W| = k$ is the probability that exactly $k$ of the elements of $S$ are in $W$ and exactly $s - k$ of the elements are not in $W$.
  This probability fits the binomial distribution with $s$ trials, $k$ successes, and probability of success $2^{-d}$.
  Thus
  \begin{equation*}
    p = \binom{s}{k} (2^{-d})^k (1 - 2^{-d})^{s - k}.
  \end{equation*}
  By finding the zeros of the first derivative of $p$ with respect to $d$, we see that $p$ achieves a maximum value at $d = -\log \frac{k}{s}$.
  \todo{Cite a text giving the first derivative of the binomial distribution.}
  When $d$ takes this value,
  \begin{equation*}
    p = \binom{s}{k} \frac{k^k}{s^s} (s - k)^{s - k}.  % so fun!
  \end{equation*}
  Expanding the binomial coefficient as $\frac{s!}{k!(s - k)!}$ and applying Stirling's approximation,
  \begin{equation*}
    \sqrt{2 \pi m} \left(\frac{m}{e}\right)^m < m! < e^{\frac{1}{12m}} \sqrt{2 \pi m} \left(\frac{m}{e}\right)^m,
  \end{equation*}
  yields the upper bound
  \begin{equation*}
    %p & < \frac{\sqrt{2 \pi s} \left(\frac{s}{e}\right)^s e^{\frac{1}{12s}} k^k (s - k)^{s - k}}{\sqrt{2 \pi (s - k)} \left(\frac{s - k}{e}\right)^{s - k} e^{\frac{1}{12 (s - k) + 1}} \sqrt{2 \pi k} \left(\frac{k}{e}\right)^k e^{\frac{1}{12 k + 1}}} \\
    %p & < \frac{\sqrt{2 \pi s} \left(\frac{s}{e}\right)^s e^{\frac{1}{12s}} k^k (s - k)^{s - k}}{\sqrt{2 \pi (s - k)} \left(\frac{s - k}{e}\right)^{s - k} \sqrt{2 \pi k} \left(\frac{k}{e}\right)^k} \\
    p < \frac{e^{\frac{1}{12s}}}{\sqrt{2 \pi}} \left(\frac{s}{k (s - k)}\right)^{\frac{1}{2}}. \qedhere
  \end{equation*}
  %\todo{See \url{http://faculty.cs.tamu.edu/klappi/random-s06/useful_bounds.pdf} for better bounds, perhaps.}
\end{proof}

Under what relationship between $k$ and $|S|$ does the probability become at most one half?

\begin{lemma}\label{lem:quad}
  Let $n$, $k$, $d$, $S$, and $s$ be as in the hypothesis of \autoref{lem:prob}.
  If $s \geq 4$ and $k$ satisfies either of the inequalities
  \begin{equation*}
    k \leq \frac{1}{2}\left(s - \frac{1}{\alpha} \sqrt{s (s - 4)}\right)
  \end{equation*}
  or
  \begin{equation*}
    k \geq \frac{1}{2}\left(s + \frac{1}{\alpha} \sqrt{s (s - 4)}\right),
  \end{equation*}
  where $\alpha = \frac{20}{103}$, then
  \begin{equation*}
    \frac{e^{\frac{1}{12s}}}{\sqrt{2 \pi}} \left(\frac{s}{k (s - k)}\right)^{\frac{1}{2}} \leq \frac{1}{2}.
  \end{equation*}
\end{lemma}
\begin{proof}
  Since $s \geq 4$, we have $\frac{1}{12 s} \leq \frac{1}{48}$, so $e^\frac{1}{12 s} \leq e^{\frac{1}{48}} \leq 1.03$.
  Also, the value of $\frac{1}{\sqrt{2 \pi}}$ is less than $0.4$.
  Thus it suffices to show that
  \begin{equation*}
    \frac{1.03}{0.4} \left(\frac{s}{k (s - k)}\right)^{\frac{1}{2}} \leq \frac{1}{2},
  \end{equation*}
  or equivalently,
  \begin{equation*}
    \frac{s}{k (s - k)} \leq \left(\frac{20}{103}\right)^2 = \alpha^2.
  \end{equation*}

  We can rearrange this inequality so that it looks like the more familiar quadratic inequality in the variable $k$,
  \begin{equation*}
    -\alpha^2 k^2 + \alpha^2 k s - s \geq 0.
  \end{equation*}
  The discriminant of this quadratic polynomial is $\alpha^2 s^2 - 4(-\alpha^2)(-s)$, which equals $\alpha^2 s (s - 4)$.
  Since $s \geq 4$ by hypothesis, the polynomial has real roots.
  Since this polynomial is convex, the inequality is satisfied when $k$ lies outside the open interval whose endpoints are the two (possibly equal) roots.
  The quadratic formula proves that the roots of the polynomial are
  \begin{equation*}
    \frac{s}{2} \pm \frac{1}{2 \alpha^2} \sqrt{\alpha^2 s (s - 4)}.
  \end{equation*}
  We conclude that the upper bound in the conclusion of the lemma is satisfied if $k$ is outside the interval
  \begin{equation*}
    \left(\frac{1}{2}\left(s - \frac{1}{\alpha} \sqrt{s (s - 4)}\right), \frac{1}{2}\left(s + \frac{1}{\alpha} \sqrt{s (s - 4)}\right)\right). \qedhere
  \end{equation*}
\end{proof}

If the probability of a Bernoulli trial is less than one half, this provides an upper bound on the binary entropy of the probability.

\begin{lemma}
  Let $n$, $k$, $d$, $S$, and $s$ be as in the hypothesis of \autoref{lem:quad}.
  Suppose $W \gets \mc{W}^*_n$ and let $A$ denote the Bernoulli random variable for the event $|S \cap W| = 1$, conditioned on the event $D = d$.
  We have
  \begin{equation*}
    H_b(A) \leq -2 \frac{e^{\frac{1}{12s}}}{\sqrt{2 \pi}} \left(\frac{s}{k (s - k)}\right)^{\frac{1}{2}} \log \left(\frac{e^{\frac{1}{12s}}}{\sqrt{2 \pi}} \left(\frac{s}{k (s - k)}\right)^{\frac{1}{2}}\right).
  \end{equation*}
\end{lemma}
\begin{proof}
  If $p$ is at most $q$ and $q \leq \frac{1}{2}$, then $H_b(p) \leq -2 q \log q$.
  Apply this fact with the previous two lemmas, choosing $q$ to be $\frac{e^{\frac{1}{12s}}}{\sqrt{2 \pi}} \left(\frac{s}{k (s - k)}\right)^{\frac{1}{2}}$.
\end{proof}

\begin{lemma}
  Let $n$, $d$, $S$, and $s$ be as in the hypothesis of the previous lemma.
  For any $k \in [2^n]$, if
  \begin{equation*}
    p^{= k} = \Pr_{W \gets \mc{W}_n^*}[ |S \cap W| = k | D = d]
  \end{equation*}
  and
  \begin{equation*}
    p^{\geq k} = \Pr_{W \gets \mc{W}_n^*}[ |S \cap W| \geq k | D = d],
  \end{equation*}
  then
  \begin{equation*}
    p^{\geq k} = \sum_{i = k}^{s} p^{= i}.
  \end{equation*}
\end{lemma}
\begin{proof}
  The first equality holds because
  \begin{align*}
    p^{\geq k} & = \Pr_{W \gets \mc{W}_n^*}\left[ |S \cap W| \geq k \middle| D = d \right] \\
    & = \Pr_{W \gets \mc{W}_n^*}\left[ \bigcup_{i = k}^{s} |S \cap W| = i \middle| D = d \right] \\
    & = \sum_{i = k}^{s} \Pr_{W \gets \mc{W}_n^*}\left[ |S \cap W| = i \middle| D = d\right] \\
    & = \sum_{i = k}^{s} p^{= i},
  \end{align*}
  where the third equality holds because the events $|S \cap W| = i$ and $|S \cap W| = j$ are mutually exclusive for any distinct $i$ and $j$ in $[2^n]$.
\end{proof}

\section{Todo}

Does the lower bound for adaptive $\mc{Q}_1$ witness-finding algorithms with high probability of success (\autoref{cor:isolationalg}) hold for any $\mc{Q}_k$ witness-finding algorithm, for any fixed positive integer $k$?
(It certainly doesn't hold when $k = 0$, because of \autoref{thm:naive}.)
This question may be answered by adapting the proof by probabilistic method given in \autocite{ablp91}.

\printbibliography

\end{document}
